Gender Segregation in Occupations: Preferences or Homophily?

### 2019-08-04 paper/Paper_2019_July_v1.tex
rewritten through the data section

%\usepackage[margin = 1.5in]{geometry}

@@

%\newpage
%\tableofcontents

@@

Section \ref{model} introduces the structure of the transferable utility matching model. Section \ref{equilibrium} describes equilibrium wages. Section \ref{empirical} describes the empirical specification and first stage of estimation, recovering the firm side parameters and reservation wages. Section \ref{stage2} describes the second stage of estimation, the decomposition of worker utility taking reservation wages as given. Section \ref{data} describes the data sources. Section \ref{results} presents results, and Section \ref{counterfactuals} presents simulations of segregation dynamics, including counterfactuals.

\section{Matching Model with Observed Transfers} \label{model}

The identification strategy in this paper builds off empirical applications of matching models to marriage markets (e.g. \citeA{Choo2006} and \citeA{Chiappori2015}), but differs in that I use data on transfers (wages) to separately identify worker and firm payoffs. In the marriage market, a lack of data on transfers means that only the sum of wage and non-wage utility is identified.\footnote{\citeA{Fox2008c} is able to estimate payoffs from the interaction of characteristics of both sides of the market but not full payoff functions.}. Because my research question is how non-wage utility varies with occupation fraction female, it is critical that I separately identify non-wage utility from utility from wages.

@@

The first step towards separate identification is to assume that observed wages are the transfers between job and worker that clear the market. Although some transfers in the labor market may be non-wage, such as enhanced benefits, wages are a natural first-order approximation. Previous literature applying transferable utility to the marriage market has assumed a logit assumption for unobserved transfers because logit is most tractable. I model wages as lognormal in order to match the observed wage distribution, which is greater than zero and has a long right tail.\footnote{Unlike the logit case, the variance of the distribution of the maximum of lognormals depends on other model parameters to be estimated, so the variance needs to be jointly estimated unlike in \citeA{Salanie2014a} for example.} Second, I define an unmatched worker to be a worker who is unemployed or out of the labor force, and an unmatched job to be a posted vacancy that has remained unfilled for a period of time. Since vacancy data is potentially a poor proxy for what it means for a job to be unmatched, I try to limit the importance of the vacancy data in my estimation procedure by choosing a specification that in principle does not require it, by assuming non-negative profit on the job side.

@@

\subsection{Payoff Functions}
\subsection{Job}

@@

Jobs care only about the relative productivity of men and women and the wage that they will have to pay to an individual worker. Let $WTP^g_o$ be the willingness-to-pay of a job in occupation $o$ of hiring a worker of gender $g$. The total payoff to job $j$, $\pi^g_j$, is specified as the ratio of willingness-to-pay to the cost of hiring a worker.

@@

The ratio in equation \ref{firm} is used as opposed to an additive profit function for two reasons. First, since firms in the model only fill one vacancy at a time, rate of return is a reasonable object for firms to care about that does not require making assumptions about the number of total vacancies at a firm. By allowing the firm to maximize return on a single vacancy at a time I make a reasonable job-level approximation of a firm level profit function. Second, it is common in labor economics to specify wages as lognormally distributed. The multiplicative specification allows model wages to be normal in the estimation stage, when wages are assumed to be lognormal. The log payoff to the firm $j$ is:

@@

\subsection{Worker}
%In a competitive equilibrium firms will produce at the production level that minimizes average cost, or equivalently maximizes willingness-to-pay over cost.

Workers choose an occupation at the beginning of their lives, based on their individual tastes for occupations and the wage offer for their gender in that occupation. The job choice is binding for life.\footnote{Since there is no switching jobs, I assume that education choice is tied to job choice and do not treat education as a separate decision.} I assume that workers are myopic and do not anticipate future changes in the market that would lead them to want to switch occupations.\footnote{Lowering switching cost would lead to faster convergence to a fixed point in the fraction female by occupation. However, relaxing myopia would pose a challenge to tractability because the endogenous attribute (fraction female) would then depend on expectations over the future behavior of all agents, complicating today's choice set considerably.}

@@

The payoff to worker $i$ is specified as:
\begin{align} \label{worker}
u^i_j &= \frac{ u^g_o*Wage^g_j *  \eta^i_o } {\xi^g_j}    
\end{align}

Non-wage and wage utility thus enter the specification multiplicatively, including the taste heterogeneity term $\eta^i_o$. Define log utility to be $\bar{u}^{i}_j$ and re-parameterizing in terms of logs we have:

@@

Workers choose a job $j$ to maximize log utility $\bar{u}^{i}_j$. The common taste parameter for non-employment $\bar{u}^g_N$, where non-employment means either not in the labor force or unemployed, is normalized to zero. It is also assumed that in non-employment workers receive no wages or value from job amenities. The payoff to non-employment ($\bar{u}^i_N$) is therefore  $ \bar{\eta}^i_N $, the idiosyncratic taste for non-employment.

@@

Furthermore worker utility will be decomposed into a common component and a dependence on the fraction female in the occupation, which is the endogenous amenity of interest. Let $F_o$ be the fraction female in occupation $o$. To avoid simultaneity, the fraction female is defined as the fraction female of the occupation in period $t-1$, that is, the fraction female that would be observed by the worker as they make their occupation decisions.

\begin{align*}
\bar{u}^{i}_j &= \bar{u}^{g}_o + \widebar{Wage}^g_j   + \bar{\eta}^i_o - \bar{\xi}^g_j \\
    &= \alpha^g_o +   \gamma^g F_o + \widebar{Wage}^g_j   + \bar{\eta}^i_o - \bar{\xi}^g_j
\end{align*}
% \frac{\beta^g}{\sigma^g_{\eta}}  X_o+

@@

% REPLACE
The lifetime income distributions by gender and occupation needed to clear the market are determined in equilibrium to equate demand and supply, and are made up of two components: a job-specific component, and an aggregate component that varies by occupation and gender. The job-specific component reflects the idiosyncratic disutility of taking a specific job. In equilibrium firms must make workers indifferent across individual jobs within an occupation, and therefore firms perfectly compensate workers for idiosyncratic disutility. Wage heterogeneity within occupation and gender emerges from these job-specific compensating differentials.

In addition to an individual job level compensating differential, there is a component of wages which is common across individuals within gender and occupation, and serves to equate supply and demand. It is a function of all parameters in the model, specifically, both of the utility that workers receive from occupations as well as the value that jobs have for male and female workers. This component of wage can be thought of as the result of a market-wide open-ended ascending price auction where jobs make wage bids for workers, but each job can only ``win" one worker, and there are no search frictions. The assumption of no search frictions is made more reasonable by the fact that I am considering lifetime occupation choices, not individual jobs. In order for equilibrium wages to not depend on sample size I must assume a large number of workers and jobs of each type \cite{Galichon2013b}. In my model this means a large number of men and women and jobs in each occupation.\footnote{For reference, simulations of an ascending price auction with sample size of 20 already produce a wage vector similar to the equilibrium vector assumed by the large sample. I perform the simulations using a modified version of the auction mechanism outlined in \citeA{RothSotomayor} page 209.}

The equilibrium wage vector I propose,  $\widebar{Wage}^g_j$, equates supply and demand for workers by gender and jobs by occupation. It does so by compensating the worker for the idiosyncratic dis-amenities of a particular job within occupation, and through a common component $W^g_o$, which varies by gender and occupation.

@@

A firm receives a draw of $\xi^M_j$ and $\xi^F_j$, which is the amenity value of the job to men and women respectively. The firm then chooses to hire a male or female worker based on the overall productivity of men and women in that occupation ($WTP^g_o$) and the cost of hiring men and women which varies in order to compensate exactly for the utility or disutility experienced by the worker due to $\xi^M_j$ and $\xi^F_j$. This is a compensating differential at the job level. Compensating differentials also emerge at the occupation level through the common component of wages, $W^g_o$.

The fact that jobs care only about whether to hire a male or female worker and the wage that they must pay means that conditional on gender, all workers are equally productive. This assumption is important because it allows me to separate the matching problem into two separate discrete choice problems, one for each side of the market \cite{Galichon2013}. The counterfactual dynamics I am interested in are at the level of occupation and gender, and productivity and utility are allowed to vary freely by occupation and gender. Therefore including individual worker productivity heterogeneity should not have a major impact on my results. In terms of biasing parameter estimates, I may overestimate how undesirable small occupations are. High wages in small occupations may be due to selecting the most skilled workers rather than dis-amenities. Similarly I may overestimate the value of amenities in large occupations if wages are low due to a lower than average level of worker skill, rather than attractive amenities.
% JUSTIFY MORE WHY ORTHOGONAL

The worker utility function therefore takes into account the job dis-amenities $\xi^M_j$ and $\xi^F_j$ as well as the workers taste for that occupation, and because $log(Wage^g_j) = log(W^g_o * \xi^g_j) = \bar{W^g_o} + \bar{\xi^g_j}$, workers are exactly compensated for $ \xi^g_j$ in the utility function. Plugging in and cancelling we get equilibrium utility for workers of

\begin{align} \label{utility}
\bar{u}^i_o &= \bar{u}^{g}_o + \widebar{Wage}^g_j   + \bar{\eta}^i_o - \bar{\xi}^g_j  \\ \nonumber
&= \bar{u}^{g}_o +  \bar{W}^g_o + \bar{\xi}^g_j  + \bar{\eta}^i_o - \bar{\xi}^g_j  \\ \nonumber
&=  \bar{u}^{g}_o + \bar{W}^g_o  + \bar{\eta}^i_o  \\ \nonumber
&=  \alpha^g_o + \gamma^g F_o + \bar{W}^g_o  + \bar{\eta}^i_o 
\end{align}

The equilibrium payoff to a firm is then

@@

The fact that this wage vector supports the pairwise stable matching follows from the optimization problems of jobs and workers. See appendix \ref{sec.Stability} for proof of stability.

@@

\subsection{Estimation}

Estimation then follows from distributional assumptions on the unobserved heterogeneity on each side of the market. The equilibrium solution allows separate estimation of each side.

\begin{assumption}
Let the heterogeneity in job amenities, $\xi^g_j$, be distributed lognormal, and independently across $j$, such that $\bar{\xi}^g_j = ln(\xi^g_j)$ is distributed normally with location parameter normalized to zero and scale parameters for each gender, $\sigma^F_{\xi}$ and $\sigma^M_{\xi}$.
\end{assumption}

Since we observe log wages $\widebar{Wage}^g_j$, the expectation of observed wages suggests itself as an estimator for $\bar{W}^g_o$. Unfortunately there is a wrinkle to this strategy. There is a selection problem in that the only wages that are observed are the wages that maximize the job's choice over male, female, or not hiring any worker. The choice to not hire at all is especially problematic since it implies a truncation point that depends on $W^g_o$ which is unknown. Likewise, the selection effects depend on the scale of the unobserved job heterogeneity, $\sigma^g_\xi $, which is unknown.\footnote{Estimating this parameter is important because is governs the elasticity of labor supply to changes in wages or other non-wage amenities such as the gender ratio in the counterfactual, as well as the demand response to changes in equilibrium wage. Unlike \citeA{Chiappori2015}, who uses many markets to identify these parameters, I use the observed wage distribution as suggested in \citeA{Salanie2014a}.} The problem therefore suggests joint likelihood estimation with selection and truncation, where the selection takes the form of a Tobit Type 5, to estimate $WTP^g_o$, $W^g_o$ and $\sigma^g_{\xi}$. See Appendix \ref{sec.likelihood} for the likelihood function in the notation of \citeA{Amemiya1985} and my notation.

@@

A firm will only hire a worker if the willingness-to-pay for that worker is higher than the wage the worker will accept in equilibrium, which means the payoff $ \pi^g_j $ must be greater than one.

$$ \pi^g_j = \frac{WTP^g_o}{Wage^g_j} \geq 1 $$

\begin{assumption}
Jobs will remain unfilled if the wage is higher than the willingness-to-pay for both men and women.
\end{assumption}

$$ \pi^F_j = \frac{WTP^F_o}{Wage^F_j} < 1 \hspace{3mm} \text{and}  \hspace{3mm} \pi^M_j = \frac{WTP^M_o}{Wage^M_j} < 1 \hspace{3mm} \implies \hspace{3mm} \text{$j$ unfilled}$$

@@

The above assumption implies that jobs are unfilled when both $\bar{\xi}^M_j$ and $\bar{\xi}^F_j$ are very high. High $\bar{\xi}^M_j$ and $\bar{\xi}^F_j$ means very high dis-amenity value to both men and women, requiring high wages to compensate. So if a job is unattractive enough, it will not be filled. I make the assumption of a fixed outside option because then the model is identified without use of data on vacancies, allowing me to test robustness to the use of the vacancy data. I can estimate the willingness-to-way of the firms with the maximum observed wage by gender and occupation. This strategy is less stable than using vacancy data.

@@

The location parameter of $\eta^i_o$ is normalized to zero, and the scale parameter is estimated separately for each gender, resulting in two scale parameters $\sigma^M_{\eta}$ and $\sigma^F_{\eta}$ for men and women respectively. In \citeA{Chiappori2015} the scale parameters require many markets to estimate but with observed transfers we can rely on variation in lifetime income across occupations to trace out the scale.

I leverage well known properties of extreme value distributions to obtain choice probabilities, or the probability that the utility from one option is higher than the utility of all other options. Recall that $log(u^i_g) = log(u^g_o) + log(W^g_o)  + \eta^i_o \equiv  \bar{u}^{g}_o + \bar{W}^g_o   + \eta^i_o  $.

@@

$$ Pr(i \in g \text{ chooses } \forall  j \in o) = \frac{exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})}{\sum_o exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})}$$

\begin{assumption}
Let $u^i_N = \eta^i_N$. This implies that the $u^g_N=0$ and no wages are received, leaving only idiosyncratic taste for non-employment $\eta^i_N$.
\end{assumption}

@@

We can then define the probability of choosing occupation $o$ relative to the probability of choosing non-employment.

$$ \frac{Pr(i \in g \text{ chooses } \forall j \in o)}{ Pr(i \in g \text{ chooses }N)}= \frac{ \frac{exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})}{\sum_Y exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})}  } {\frac{1}{\sum_Y exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})} } = exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})$$

Let $s^g_o$ be the share of workers of gender $g$ who match to occupation $o$, and $s^g_N$ the share that choose non-employment. These shares are observed. Rearranging from above we have

%$$ \Pi^g_o = \sigma^g_{\xi} ln(  \mu_{X|Y}) - \sigma^g_{\xi}  ln(\mu_{0|Y})= \sigma^g_{\xi} ln(  \mu^g_o) -  \sigma^g_{\xi} ln(\mu_{0Y}) $$
\begin{align} \label{logit}
  \sigma^g_{\eta} ln(s^g_o) -  \sigma^g_{\eta} ln(s^g_N) &= \bar{u}^{g}_o + \bar{W}^g_o \\
     ln(s^g_o) - ln(s^g_N) &= \frac{\bar{u}^{g}_o + \bar{W}^g_o}{ \sigma^g_{\eta} } \\
     &= \frac{\alpha^{g}_o + \gamma^g F_o + \bar{W}^g_o}{ \sigma^g_{\eta} }
\end{align}

%\section{Preference for Gender Ratio} \label{stage2}

Recall that in a logit model the mean utility in occupation $o$ is equal to the log of the ratio of the share employed ($s^g_o$) in that occupation to the share in non-employment ($s^g_N$), or $ ln(s^g_o) - ln(s^g_N) $. The mean utility can then be decomposed into wage and non-wage utility which are common to all worker and job matches of a given occupation and gender. The non-wage component is $\bar{u}^{g}_o$ and the wage component $\bar{W}^g_o$. The impact of both on mean utility is mediated by the scale of the taste heterogeneity ($\sigma^g_{\eta}$).

\begin{align*}
     ln(s^g_o) - ln(s^g_N) = \frac{\bar{u}^{g}_o + \bar{W}^g_o}{ \sigma^g_{\eta} } = \frac{\bar{\alpha}^{g}_o + \gamma^g F_o  + \bar{W}^g_o}{ \sigma^g_{\eta} } 
\end{align*}

This equation cannot be estimated in the cross section because we have one $\alpha^g_o$ parameter for each occupation, in addition to $\gamma^g_o$ and $\sigma^g_{\eta}$, and only $n_o$ moments.

In the sections below I introduce several strategies to identify the parameters $\frac{\gamma^g}{\sigma^g_{\eta}}$ and $\frac{1}{\sigma^g_{\eta}}$. First, I pool all six cross sections of data (1960-2012), allowing me to estimate the occupation-specific intercepts, $\alpha^g_o$ and use time variation in fraction female and reservation wages to identify coefficients $\frac{\gamma^g}{\sigma^g_{\eta}}$ and $\frac{1}{\sigma^g_{\eta}}$. Unfortunately using time variation means that estimates of $\frac{\gamma^g}{\sigma^g_{\eta}}$ and $\frac{1}{\sigma^g_{\eta}}$ likely suffer from omitted variable bias. Changes over time in both the gender ratio and the reservation wage are likely correlated with changes in unobserved occupation attributes.

>>>>

\usepackage[margin = 1in]{geometry}

@@

\newpage
\tableofcontents

@@

Section \ref{model} introduces the structure of the transferable utility matching model and equilibrium wages. Section \ref{empirical} describes the empirical specification including recovering the firm side parameters and wage offer distribution, and the decomposition of worker utility into fixed occupation attributes and gender preference. Section \ref{data} describes the data sources. Section \ref{results} presents results, and Section \ref{counterfactuals} presents simulations of segregation dynamics, including counterfactuals.

\section{Model} \label{model}

The identification strategy in this paper builds off empirical applications of matching models to marriage markets (e.g. \citeA{Choo2006} and \citeA{Chiappori2015}), but differs in that I use data on transfers (wages) to separately identify worker and firm payoffs. In the marriage market, a lack of data on transfers means that only the sum of wage and non-wage utility is identified.\footnote{\citeA{Fox2008c} is able to estimate payoffs from the interaction of characteristics of both sides of the market but not full payoff functions.}. Another key difference from previous literature is that I use the distribution of transfers (wages) for identification.

%Because my research question is how non-wage utility varies with occupation fraction female, it is critical that I separately identify non-wage utility from utility from wages.

@@

The first step towards separate identification of firm and worker non-wage parameters is to assume that observed wages are the transfers between job and worker that clear the market. Although some transfers in the labor market may be non-wage, such as enhanced benefits, wages are a natural first-order approximation. Previous literature applying transferable utility to the marriage market has assumed a logit assumption for unobserved transfers because logit is most tractable. I model wages as lognormal in order to match the observed wage distribution, which is greater than zero and has a long right tail.\footnote{Unlike the logit case, the variance of the distribution of the maximum of lognormals depends on other model parameters to be estimated, so the variance needs to be jointly estimated unlike in \citeA{Salanie2014a} for example.} Second, I define an unmatched worker to be a worker who is unemployed or out of the labor force, and an unmatched job to be a posted vacancy that has remained unfilled for a period of time. Since vacancy data is potentially a poor proxy for what it means for a job to be unmatched, I try to limit the importance of the vacancy data in my estimation procedure by choosing a specification that in principle does not require it, by assuming non-negative profit on the job side. Thus identification relies more heavily on the wage distribution.

@@

\subsection{Payoff Functions}
\subsubsection{Firm}

@@

% the relative productivity of men and women and the wage that they will have to pay to an individual worker.

Firms are willing to pay different amounts for male and female workers. I do not take a stance as to the cause of the difference. Let $WTP^g_o$ be the most that a job in occupation $o$ would be willing to pay to hire a worker of gender $g$. The total payoff to job $j$, $\pi^g_j$, also depends on the cost of hiring a worker, ${Wage^g_j}$, which varies at the job level according to how attractive the job is to workers.

@@

I specify the payoff function as a ratio for both economic and computational reasons. First, maximizing rate of return is consistent with a firm only filling one vacancy at a time, and allows me to avoid making assumptions about the number of vacancies at a firm or complementarities between those vacancies. Second, the multiplicative specification allows wages to be lognormally distributed, which is common in labor economics to match the observed shape of the wage distribution, while still additive and normal when logged, which is computationally attractive.

The log payoff to the firm $j$ is:

%First, since firms in the model only fill one vacancy at a time, rate of return is a reasonable object for firms to care about that does not require making assumptions about the number of total vacancies at a firm. By allowing the firm to maximize return on a single vacancy at a time I make a reasonable job-level approximation of a firm level profit function.

@@

\subsubsection{Worker}
%In a competitive equilibrium firms will produce at the production level that minimizes average cost, or equivalently maximizes willingness-to-pay over cost.

Workers choose an occupation at the beginning of their lives, based on their individual tastes for occupations and the wage offer for their gender in that occupation. The job choice is binding for life.\footnote{Since there is no switching jobs, I assume that education choice is tied to job choice and do not treat education as a separate decision. Lowering switching cost would lead to faster convergence to a fixed point in the fraction female by occupation.} I assume that workers are myopic and do not anticipate future changes in the market that would lead them to want to switch occupations.\footnote{Relaxing myopia would mean that the endogenous attribute of interest, the fraction female, would depend on expectations over other workers' occupation choices, which in turn are affected by the fraction female. This would pose a challenge to tractability.}

@@

%Non-employment is defined as either not in the labor force or unemployed, and the payoff to non-employment is $ \bar{\eta}^i_N $, the idiosyncratic taste for non-employment.

The payoff to worker $i$ from job $j$ is specified as:
\begin{align} \label{worker}
u^i_j &= \frac{ u^g_o*Wage^g_j *  \eta^i_o } {\xi^g_j}    
\end{align}

Defining log utility to be $\bar{u}^{i}_j$ and re-parameterizing in terms of logs we have:

@@

%Workers choose a job $j$ to maximize log utility $\bar{u}^{i}_j$. The common taste parameter for non-employment $\bar{u}^g_N$, where non-employment means either not in the labor force or unemployed, is normalized to zero. It is also assumed that in non-employment workers receive no wages or value from job amenities. The payoff to non-employment ($\bar{u}^i_N$) is therefore  $ \bar{\eta}^i_N $, the idiosyncratic taste for non-employment.

@@

Furthermore worker utility will be decomposed into a common component and a dependence on the fraction female in the occupation, which is the endogenous amenity of interest. Let $F_o$ be the fraction female in occupation $o$. To avoid simultaneity, the fraction female is defined as the fraction female of the occupation among older workers (ages 35-65), that is, the fraction female that would be observed by the worker as they make their occupation decisions. To make the timing assumption clear I use subscript $t-1$ for the fraction female.

\begin{align*}
\bar{u}^{i}_j &= \bar{u}^{g}_o + \widebar{Wage}^g_j   + \bar{\eta}^i_o - \bar{\xi}^g_j \\
    &= \alpha^g_o +   \gamma^g F_{o,t-1} + \widebar{Wage}^g_j   + \bar{\eta}^i_o - \bar{\xi}^g_j
\end{align*}
% \frac{\beta^g}{\sigma^g_{\eta}}  X_o+

@@

% REPLACE
%The lifetime income distributions by gender and occupation needed to clear the market are determined in equilibrium to equate demand and supply, and are made up of two components: a job-specific component, and an aggregate component that varies by occupation and gender. The job-specific component reflects the idiosyncratic disutility of taking a specific job. In equilibrium firms must make workers indifferent across individual jobs within an occupation, and therefore firms perfectly compensate workers for idiosyncratic disutility. Wage heterogeneity within occupation and gender emerges from these job-specific compensating differentials.

%In addition to an individual job level compensating differential, there is a component of wages which is common across individuals within gender and occupation, and serves to equate supply and demand. It is a function of all parameters in the model, specifically, both of the utility that workers receive from occupations as well as the value that jobs have for male and female workers.

%\footnote{For reference, simulations of an ascending price auction with sample size of 20 already produce a wage vector similar to the equilibrium vector assumed by the large sample. I perform the simulations using a modified version of the auction mechanism outlined in \citeA{RothSotomayor} page 209.}

The equilibrium wages I propose, $\widebar{Wage}^g_j$, equate supply and demand for workers by gender and jobs by occupation. They do so by compensating the worker for the idiosyncratic dis-amenities of a particular job within occupation, and through a common component $W^g_o$, which varies by gender and occupation. These equilibrium wages support the pairwise stable matching of workers and jobs through the optimization on both sides of the market. See appendix \ref{sec.Stability} for proof of stability.

@@

A firm receives a draw of $\xi^M_j$ and $\xi^F_j$, which is the amenity value of the job to men and women respectively. The firm then chooses to hire a male or female worker based on the overall productivity of men and women in that occupation ($WTP^g_o$) and the cost of hiring men and women which varies in order to compensate exactly for the utility or disutility experienced by the worker due to $\xi^M_j$ and $\xi^F_j$. This is a compensating differential at the job level.

Compensating differentials also emerge at the occupation level through the common component of wages, $W^g_o$, which will vary according to the utility that workers receive from occupations, as well as the value that jobs have for male and female workers, in order to equate supply and demand. This component of wages can be thought of as the result of a market-wide open-ended ascending price auction where jobs make wage bids for workers, but each job can only ``win" one worker.\footnote{In order for equilibrium wages to not depend on sample size I must assume a large number of workers of each gender and jobs of each occupation \cite{Galichon2013b}.}

% , and there are no search frictions. The assumption of no search frictions is made more reasonable by the fact that I am considering lifetime occupation choices, not individual jobs.

Plugging in equilibrium wages we have the following equilibrium utility for workers:

\begin{align} \label{utility}
\bar{u}^i_o &= \bar{u}^{g}_o + \widebar{Wage}^g_j   + \bar{\eta}^i_o - \bar{\xi}^g_j  \\ \nonumber
&= \bar{u}^{g}_o +  \bar{W}^g_o + \bar{\xi}^g_j  + \bar{\eta}^i_o - \bar{\xi}^g_j  \\ \nonumber
&=  \bar{u}^{g}_o + \bar{W}^g_o  + \bar{\eta}^i_o  \\ \nonumber
&=  \alpha^g_o + \gamma^g F_{o,t-1} + \bar{W}^g_o  + \bar{\eta}^i_o 
\end{align}

Workers are exactly compensated for the job dis-amenities, $ \xi^g_j$, in equilibrium, leaving the taste for occupation, $\bar{\eta}^i_o$, as the heterogeneity at the individual worker level.

%The worker utility function therefore takes into account the job dis-amenities $\xi^M_j$ and $\xi^F_j$ as well as the workers taste for that occupation, and because $log(Wage^g_j) = log(W^g_o * \xi^g_j) = \bar{W^g_o} + \bar{\xi^g_j}$,

The equilibrium payoff to a firm is

@@

Firms care only about the value proposition of hiring a man or woman ($\widebar{WTP}^g_o -  \bar{W}^g_o$), and the idiosyncratic appeal of their job to men and women, $\bar{\xi}^g_j$. This means that, as on the worker side, there is only one source of heterogeneity in the equilibrium payoff. This is a critical assumption because it allows me to separate the matching problem into two separate discrete choice problems, one for each side of the market \cite{Galichon2013}. The cost of this assumption is that there is no individual worker productivity heterogeneity. Selection on productivity is orthogonal to my research question, but it may lead me to underestimate the value of amenities of small occupations and overestimate the appeal of large occupations.

%The counterfactual dynamics I am interested in are at the level of occupation and gender, and productivity and utility are allowed to vary freely by occupation and gender. Therefore including individual worker productivity heterogeneity should not have a major impact on my results. In terms of biasing parameter estimates, I may overestimate how undesirable small occupations are. High wages in small occupations may be due to selecting the most skilled workers rather than dis-amenities. Similarly I may overestimate the value of amenities in large occupations if wages are low due to a lower than average level of worker skill, rather than attractive amenities.
% JUSTIFY MORE WHY ORTHOGONAL

@@

Estimation follows from distributional assumptions on the unobserved heterogeneity on each side of the market.

\subsection{Firm Estimation}

\begin{assumption}
Let the heterogeneity in job amenities for each gender, $\xi^g_j$, be distributed lognormal and independent across $j$, such that $\bar{\xi}^g_j = ln(\xi^g_j)$ is distributed $\mathcal{N}(0, \sigma^g_{\xi}$).
\end{assumption}

I also assume that if a job is unattractive enough, it will not be filled. This allows me to identify the model without use of vacancy data on the firm side. In practice I use vacancy data in estimation for stability of the estimator and to avoid relying too much on extreme data points for identification.

%I make the assumption of a fixed outside option because then the model is identified without use of data on vacancies, allowing me to test robustness to the use of the vacancy data. I can estimate the willingness-to-way of the firms with the maximum observed wage by gender and occupation. This strategy is less stable than using vacancy data.

@@

%A firm will only hire a worker if the willingness-to-pay for that worker is higher than the wage the worker will accept in equilibrium, which means the payoff $ \pi^g_j $ must be greater than one.

%$$ \pi^g_j = \frac{WTP^g_o}{Wage^g_j} \geq 1 $$

\begin{assumption}
A firm will not hire a worker if the wage for men is higher than the willingness-to-pay for men ($WTP^M_o$) and the wage for women is higher than the firm's willingness-to-pay for women ($WTP^F_o$).
\end{assumption}

$$ \bar{\pi}^g_j = \widebar{WTP}^g_o - \widebar{Wage}^g_j <  0 \hspace{3mm} \text{for} \hspace{3mm} g \in \{M,F\}  \hspace{3mm} \implies \hspace{3mm} \text{$j$ unfilled}$$

@@

%The above assumption implies that jobs are unfilled when both $\bar{\xi}^M_j$ and $\bar{\xi}^F_j$ are very high. High $\bar{\xi}^M_j$ and $\bar{\xi}^F_j$ means very high dis-amenity value to both men and women, requiring high wages to compensate.

We observe log wages ($\widebar{Wage}^g_j$), but the only wages that are observed are the wages that maximize the job's choice over male, female, or not hiring any worker, so the observed data are the result of both selection and truncation. The selection and truncation depends on the $\bar{W}^g_o$, which is an unknown equilibrium object to be estimated, as well as the scale of the unobserved job heterogeneity, $\sigma^g_\xi $, which is unknown.\footnote{Estimating scale is important because is governs the elasticity of labor supply to changes in wages or other non-wage amenities such as the gender ratio in the counterfactual, as well as the demand response to changes in equilibrium wage. Unlike \citeA{Chiappori2015}, who uses many markets to identify these parameters, I use the observed wage distribution as suggested in \citeA{Salanie2014a}.} To back out the unobserved wage offer distribution from the observed wages I use Tobit Type 5 maximum likelihood estimation. This allows me to estimate $WTP^g_o$, $W^g_o$ and $\sigma^g_{\xi}$ while accounting for selection and truncation. See Appendix \ref{sec.likelihood} for the likelihood function in the notation of \citeA{Amemiya1985} and my notation.

%Since we observe log wages $\widebar{Wage}^g_j$, the expectation of observed wages suggests itself as an estimator for $\bar{W}^g_o$. Unfortunately there is a wrinkle to this strategy. There is a selection problem in that .

\subsection{Worker Estimation}

@@

The location parameter of $\eta^i_o$ is normalized to zero, and the scale parameter is estimated separately for each gender, resulting in two scale parameters $\sigma^M_{\eta}$ and $\sigma^F_{\eta}$ for men and women respectively. I leverage well known properties of extreme value distributions to obtain occupation choice probabilities in terms of utility parameters. Recall that $log(u^i_g) = log(u^g_o) + log(W^g_o)  + \eta^i_o \equiv  \bar{u}^{g}_o + \bar{W}^g_o   + \eta^i_o  $.

%In \citeA{Chiappori2015} the scale parameters require many markets to estimate but with observed transfers we can rely on variation in lifetime income across occupations to trace out the scale.

@@

$$ Pr(i \in g \text{ chooses } \forall  j \in o) = \frac{exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})}{\sum_{k \in O} exp(\frac{\bar{u}^{g}_k + \bar{W}^g_k}{\sigma^g_{\eta}})}$$

\begin{assumption}
I normalize the log utility from non-employment to be zero ($\bar{u}^g_N=0$). 
\end{assumption}

@@

Since no wages are received in non-employment, this leaves only the idiosyncratic taste for non-employment $\eta^i_N$. We can then define the probability of choosing occupation $o$ relative to the probability of choosing non-employment in terms of the share of workers of gender $g$ who match to occupation $o$ ($s^g_o$), and  the share that choose non-employment ($s^g_N$), which are observed. Recall that the non-wage utility ($\bar{u}^{g}_o$) can then be decomposed into utility from the occupation ($\alpha^{g}_o$) and utility from the fraction female ($\gamma^g F_{o,t-1}$).

%$$ \frac{Pr(i \in g \text{ chooses } \forall j \in o)}{ Pr(i \in g \text{ chooses }N)}= \frac{ \frac{exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})}{\sum_Y exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})}  } {\frac{1}{\sum_Y exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})} } = exp(\frac{\bar{u}^{g}_o + \bar{W}^g_o}{\sigma^g_{\eta}})$$

\begin{equation} \label{eq:1}
\begin{split}
ln \left( \frac{Pr(i \in g \text{ chooses } \forall j \in o)}{ Pr(i \in g \text{ chooses }N)} \right) &= ln(s^g_o) - ln(s^g_N) \\
 &= \frac{\bar{u}^{g}_o + \bar{W}^g_o}{ \sigma^g_{\eta} } = \frac{\alpha^{g}_o + \gamma^g F_{o,t-1}  + \bar{W}^g_o}{ \sigma^g_{\eta} } 
 \end{split}
\end{equation}

This equation \ref{eq:1} cannot be estimated in the cross section because there are only as many moments as occupations, and I must estimate $\gamma^g$ and $\sigma^g_{\eta}$ in addition to an $\alpha^g_o$ for each occupation. Therefore I pool all six cross sections of data (1960-2012), allowing me to estimate the occupation-specific intercepts, $\alpha^g_o$ and use time variation in fraction female and reservation wages to identify coefficients $\frac{\gamma^g}{\sigma^g_{\eta}}$ and $\frac{1}{\sigma^g_{\eta}}$. In order to better match time variation in shares, a time effect $\beta^g_t$ is added. In addition, in the spirit of \citeA{Berry1994}, $\epsilon^g_{o,t}$ represents changes over time in the utility of workers due to changes in unobserved occupation attributes, so changes not due to movement in the fraction female or the reservation wages.

\begin{align*}
ln(s^g_{o,t}) - ln(s^g_{N,t}) =   \frac{ \beta^g_t + \alpha^g_o +  \gamma^g F_{o} +  \bar{W}^g_{o,t} + \epsilon^g_{o,t}}{\sigma^g_{\eta}}  \\
\end{align*}

%In the sections below I introduce several strategies to identify the parameters $\frac{\gamma^g}{\sigma^g_{\eta}}$ and $\frac{1}{\sigma^g_{\eta}}$.

Unfortunately using time variation means that estimates of $\frac{\gamma^g}{\sigma^g_{\eta}}$ and $\frac{1}{\sigma^g_{\eta}}$ likely suffer from omitted variable bias. Changes over time in both the gender ratio and the reservation wage are likely correlated with changes in unobserved occupation attributes $\epsilon^g_{o,t}$.

For example if an occupation is becoming more family friendly over time, and this causes more women to enter the occupation, the coefficient on fraction female will be biased upward for women. Similarly if occupation amenities deteriorate over time, this may be correlated with increases in wages, causing a downward bias on the coefficient on the reservation wage $W^g_o$.
